### 2023-03-30 scripts/manuscript/graveyard.qmd
Should probably commit before doing more edits

---
title: "text graveyard"
format: html
---

For the entire population of adults diagnosed with diabetes in the US there was a significant percent decrease of `r pullNumber(tab1, '-', 'Percent Change (95% CI)')` of people receiving at least three of the recommended practices from 2008 to 2020 (Table 1). The percent change observed among subgroups were either decreasing or showed no change (the 95% confidence intervals for percent change overlap with zero) (Table 1). The largest negative percent change was in the income poverty ratio group between 100% and 199% at `r pullNumber(tab1, '100% - 199%', 'Percent Change (95% CI)')`. There was also a `r pullNumber(tab1, '200% - 399%', 'Percent Change (95% CI)')` percent change in the income poverty ratio group between 200% and 399% and a `r pullNumber(tab1, 'Less than high school', 'Percent Change (95% CI)')` percent change seen in those with no high school diploma (Table 1). Other groups that had significant percent decreases include a `r pullNumber(tab1, '45 to 64', 'Percent Change (95% CI)')` percent change in adults aged 45 to 64, a `r pullNumber(tab1, 'Female', 'Percent Change (95% CI)')` percent change for adult females, a `r pullNumber(tab1, 'Any private', 'Percent Change (95% CI)')` percent change for adults with any private insurance, and a `r pullNumber(tab1, 'Public only', 'Percent Change (95% CI)')` percent change for adults with only public insurance.

## Trends in ≥ 3 preventive care practices received

Overall, we observed no statistically significant percent change of adults receiving three or more recommended preventive care practices between 2008 and 2019 (`r pullNumber(tab_sans_2020, '-', 'Percent Change (95% CI)', brackets = TRUE)`) (Table 1). Few groups exhibited significant changes in this trend over the same time period (Table 1). For adults age 75 and older, the percentage receiving three or more of the recommended care practices was flat from 2008 to 2014 (APC `r pullNumber(tab_sans_2020, '75+', 'APC Period 1 (95% CI)', brackets = T)`), and after 2015 was decreasing (APC `r pullNumber(tab_sans_2020, '75+', 'APC Period 2 (95% CI)', brackets = T)`). Adults with less than a high school education had a flat trend until 2012, which was followed by an APC in the second period of `r pullNumber(tab_sans_2020, 'Less than high school', 'APC Period 2 (95% CI)', brackets = F)`.

Both males and females showed decreasing percentages of adults receiving three or more recommended preventive care practices. The AAPC for females was `r pullNumber(tab_sans_2020, 'Female', 'APC Period 1 (95% CI)', brackets = F)` over the entire period, whereas the trend for males was flat until 2015, at which point it began to decrease (APC `r pullNumber(tab_sans_2020, 'Male', 'APC Period 2 (95% CI)', brackets = T)`). Among the insurance groups, there were no significant change with the exception of Medicaid, which showed an overall decline in percentage of adults receiving three or more practices (AAPC `r pullNumber(tab_sans_2020, 'Medicaid', 'APC Period 1 (95% CI)', brackets = F)`). The percentage of adults in the highest earning group receiving three or more recommended care practices increased annually by `r pullNumber(tab_sans_2020, '> 400%', 'APC Period 1 (95% CI)', brackets = F)` and was flat after 2010 (`r pullNumber(tab_sans_2020, '> 400%', 'APC Period 2 (95% CI)', brackets = T)`). Adults earning just above the poverty line that received three or more of the recommended practices changed by `r pullNumber(tab_sans_2020, '100% - 199%', 'APC Period 1 (95% CI)', brackets = F)`. Estimated proportions tended to be lower in 2020 than in 2019 for most groups, except Black/Not Hispanic, males, and the uninsured.

## Trends for each preventive care practice

@fig-exams and @fig-tests show the trends for individual preventive care practices and indicate which trends were significantly increasing or decreasing, or had segments that were significantly increasing or decreasing. Estimates plotted in @fig-exams can be found in the table in Appendix 1. Trends for all adults with diabetes receiving a foot examination  decreased starting in 2011 (@fig-exams). The percent of adults getting a flu vaccine increased until 2013 then began to decrease (@fig-tests). The shape of the trends in subgroups tend to mirror the overall trend for each practice, although the year at which trend values change may differ among these subgroups. The trend in some subgroups for each preventive practice do not always follow the shape of the overall trend, however. For example, the overall trend in the percentage of adults with diabetes receiving an eye examination with dilation was flat over 2008 to 2019. Within the income poverty ratio subgroups, only those in the highest poverty income ratio group (\> 400%) deviate from this flat trend by first increasing, then decreasing starting in 2013 (@fig-exams). Different groups exhibit different amounts of variation among the categories in each group. For example, the trends for all six practices split by sex tend to be very similar, whereas the gaps between the different insurance group trends are larger.