### 2023-05-19 scripts/manuscript/graveyard.qmd
Committing before making a big change to code

---
title: "text graveyard"
format: html
---

For the entire population of adults diagnosed with diabetes in the US there was a significant percent decrease of `r pullNumber(tab1, '-', 'Percent Change (95% CI)')` of people receiving at least three of the recommended practices from 2008 to 2020 (Table 1). The percent change observed among subgroups were either decreasing or showed no change (the 95% confidence intervals for percent change overlap with zero) (Table 1). The largest negative percent change was in the income poverty ratio group between 100% and 199% at `r pullNumber(tab1, '100% - 199%', 'Percent Change (95% CI)')`. There was also a `r pullNumber(tab1, '200% - 399%', 'Percent Change (95% CI)')` percent change in the income poverty ratio group between 200% and 399% and a `r pullNumber(tab1, 'Less than high school', 'Percent Change (95% CI)')` percent change seen in those with no high school diploma (Table 1). Other groups that had significant percent decreases include a `r pullNumber(tab1, '45 to 64', 'Percent Change (95% CI)')` percent change in adults aged 45 to 64, a `r pullNumber(tab1, 'Female', 'Percent Change (95% CI)')` percent change for adult females, a `r pullNumber(tab1, 'Any private', 'Percent Change (95% CI)')` percent change for adults with any private insurance, and a `r pullNumber(tab1, 'Public only', 'Percent Change (95% CI)')` percent change for adults with only public insurance.

## Trends in ≥ 3 preventive care practices received

Overall, we observed no statistically significant percent change of adults receiving three or more recommended preventive care practices between 2008 and 2019 (`r pullNumber(tab_sans_2020, '-', 'Percent Change (95% CI)', brackets = TRUE)`) (Table 1). Few groups exhibited significant changes in this trend over the same time period (Table 1). For adults age 75 and older, the percentage receiving three or more of the recommended care practices was flat from 2008 to 2014 (APC `r pullNumber(tab_sans_2020, '75+', 'APC Period 1 (95% CI)', brackets = T)`), and after 2015 was decreasing (APC `r pullNumber(tab_sans_2020, '75+', 'APC Period 2 (95% CI)', brackets = T)`). Adults with less than a high school education had a flat trend until 2012, which was followed by an APC in the second period of `r pullNumber(tab_sans_2020, 'Less than high school', 'APC Period 2 (95% CI)', brackets = F)`.

Both males and females showed decreasing percentages of adults receiving three or more recommended preventive care practices. The AAPC for females was `r pullNumber(tab_sans_2020, 'Female', 'APC Period 1 (95% CI)', brackets = F)` over the entire period, whereas the trend for males was flat until 2015, at which point it began to decrease (APC `r pullNumber(tab_sans_2020, 'Male', 'APC Period 2 (95% CI)', brackets = T)`). Among the insurance groups, there were no significant change with the exception of Medicaid, which showed an overall decline in percentage of adults receiving three or more practices (AAPC `r pullNumber(tab_sans_2020, 'Medicaid', 'APC Period 1 (95% CI)', brackets = F)`). The percentage of adults in the highest earning group receiving three or more recommended care practices increased annually by `r pullNumber(tab_sans_2020, '> 400%', 'APC Period 1 (95% CI)', brackets = F)` and was flat after 2010 (`r pullNumber(tab_sans_2020, '> 400%', 'APC Period 2 (95% CI)', brackets = T)`). Adults earning just above the poverty line that received three or more of the recommended practices changed by `r pullNumber(tab_sans_2020, '100% - 199%', 'APC Period 1 (95% CI)', brackets = F)`. Estimated proportions tended to be lower in 2020 than in 2019 for most groups, except Black/Not Hispanic, males, and the uninsured.

## Trends for each preventive care practice

@fig-exams and @fig-tests show the trends for individual preventive care practices and indicate which trends were significantly increasing or decreasing, or had segments that were significantly increasing or decreasing. Estimates plotted in @fig-exams can be found in the table in Appendix 1. Trends for all adults with diabetes receiving a foot examination  decreased starting in 2011 (@fig-exams). The percent of adults getting a flu vaccine increased until 2013 then began to decrease (@fig-tests). The shape of the trends in subgroups tend to mirror the overall trend for each practice, although the year at which trend values change may differ among these subgroups. The trend in some subgroups for each preventive practice do not always follow the shape of the overall trend, however. For example, the overall trend in the percentage of adults with diabetes receiving an eye examination with dilation was flat over 2008 to 2019. Within the income poverty ratio subgroups, only those in the highest poverty income ratio group (\> 400%) deviate from this flat trend by first increasing, then decreasing starting in 2013 (@fig-exams). Different groups exhibit different amounts of variation among the categories in each group. For example, the trends for all six practices split by sex tend to be very similar, whereas the gaps between the different insurance group trends are larger.

         ` ` = as.factor(` `) %>% fct_relevel('Overall',
                           'Age: 18 to 44',
                           'Age: 45 to 64',
                           'Age: 65 to 74',
                           'Age: 75+',
                           'Sex: Female',
                           'Sex: Male',
                           'Race/Ethnicity: Asian/Not Hispanic',
                           'Race/Ethnicity: Black/Not Hispanic',
                           'Race/Ethnicity: Hispanic',
                           'Race/Ethnicity: White/Not Hispanic',
                           'Highest degree earned: Greater than high school',
                           'Highest degree earned: High school',
                           'Highest degree earned: Less than high school',
                           'Insurance coverage: Medicaid',
                           'Insurance coverage: Medicare only',
                           'Insurance coverage: Private only',
                           'Insurance coverage: Uninsured',
                           'Poverty income ratio: > 400%',
                           'Poverty income ratio: 200% - 399%',
                           'Poverty income ratio: 100% - 199%',
                           'Poverty income ratio: < 100%')) %>%

# This code here changes the apc values to NA if they aren't significant, but I think
    # what I need to do is change the APC values if there are no identified joinpoints
    # across(
    #   c(apc_0, apc_95_lcl_0, apc_95_ucl_0),
    #   ~ case_when(p_value_0 > 0.05 &
    #                 p_value_1 > 0.05 & p_value_2 > 0.05 & p_value_3 > 0.05 ~ NA_real_,
    #               TRUE ~ .x)
    # ),
    # across(
    #   c(apc_1, apc_95_lcl_1, apc_95_ucl_1),
    #   ~ case_when(p_value_0 > 0.05 &
    #                 p_value_1 > 0.05 & p_value_2 > 0.05 & p_value_3 > 0.05 ~ NA_real_,
    #               TRUE ~ .x)
    # ),
    # across(
    #   c(apc_2, apc_95_lcl_2, apc_95_ucl_2),
    #   ~ case_when(p_value_0 > 0.05 &
    #                 p_value_1 > 0.05 & p_value_2 > 0.05 & p_value_3 > 0.05 ~ NA_real_,
    #               TRUE ~ .x)
    # ),
    # across(
    #   c(apc_3, apc_95_lcl_3, apc_95_ucl_3),
    #   ~ case_when(p_value_0 > 0.05 &
    #                 p_value_1 > 0.05 & p_value_2 > 0.05 & p_value_3 > 0.05 ~ NA_real_,
    #               TRUE ~ .x)
    # ),

```{r all-plots}

stats_table_plot_dat <- jp_proportion_plot_data %>%
  mutate(model = as.numeric(model)) %>% 
  filter(strata != 'Total') %>%
  filter(strata != 'Not available') %>% 
  filter(strata != 'Other Race/Not Hispanic') %>% 
  filter(strata != 'Other') %>% 
  mutate(variable_text = case_when(variable == 'a1c' ~ 'Two or more A1C tests',
                                   variable == 'CHOL_all' ~ 'Cholesterol tested',
                                   variable == 'DENT_all' ~ 'One or more dentist visits',
                                   variable == 'eye-exam' ~ 'Eye exam with dilation',
                                   variable == 'flu' ~ 'Received flu vaccine',
                                   variable == 'foot' ~ 'Foot examination'),
         strata = case_when(strata == 'stat_0' ~ 'Overall',
                            TRUE ~ strata)) %>% 
  mutate(strata = as.factor(strata)) %>% 
  split(., f = .$stratifier)

stats_table_plot_dat <- map(stats_table_plot_dat, function(.x) {
  .x %>% 
    mutate(strata = fct_drop(strata))
})

overall_data <- stats_table_plot_dat$overall %>% 
  mutate(strata = case_when(strata == 'Overall' ~ 'z'))
# Remove the overall data from the list
stats_table_plot_dat <- stats_table_plot_dat[-4]
# Relevel the poverty factors
stats_table_plot_dat$poverty$strata <-
  factor(
    stats_table_plot_dat$poverty$strata,
    levels = c('High income', 'Middle income', 'Low income', 'Poor/Negative'),
    labels = c(
      '> 400% ',
      '200% - 399%',
      '100% - 199%',
      '< 100%'

    )
  )
test_plots <- stats_table_plot_dat %>% 
   map(
    ~ ggplot(filter(.x, variable %in% c('a1c', 'CHOL_all', 'flu')),
             aes(
      x = year,
      y = p,
      group = strata,
      color = strata
    )) +
      #geom_point() +
      geom_line(aes(x = year, y = model, group = strata, color = strata)) +
      geom_line(data = filter(overall_data, variable %in% c('a1c', 'CHOL_all', 'flu')),
                  aes(x = year, y = model),
                  color = 'black',
                  linetype = 'dashed') +
      facet_wrap( ~ variable_text, ncol = 3) +
      theme_classic() +
      scale_y_continuous(
        limits = c(00, 100),
        breaks = seq(0, 100, 25),
        labels = seq(0, 100, 25)
      ) +
      scale_x_continuous(
        limits = c(2008, 2020),
        breaks = seq(2008, 2020, 2),
        labels = seq(2008, 2020, 2)
      ) +
      scale_color_brewer(type = 'qual', palette = 'Set2') +
      theme(axis.text = element_text(color = 'black'),
            axis.text.x = element_text(angle = 45, hjust = 0.9, vjust = 0.95),
            legend.title = element_text(size = 8),
            legend.text = element_text(size = 6)) +
      labs(x = 'Year', y = 'Percentage')
  )

test_plots[[1]] <- test_plots[[1]] + labs(color = 'Age (grouped)') +
  theme(axis.title.x = element_blank())
test_plots[[2]] <-
  test_plots[[2]] + labs(color = 'Highest degree \nobtained') + theme(
    strip.background =  element_blank(),
    strip.text = element_blank(),
    axis.title.x = element_blank()
  )
test_plots[[3]] <-
  test_plots[[3]] + labs(color = "Insurance") + theme(
    strip.background =  element_blank(),
    strip.text = element_blank(),
    axis.title.x = element_blank()
  )
test_plots[[4]] <-
  test_plots[[4]] + labs(color = 'Poverty income ratio') + theme(
    strip.background =  element_blank(),
    strip.text = element_blank()
  )
test_plots[[5]] <-
  test_plots[[5]] + labs(color = 'Race & Ethnicity') + theme(
    strip.background =  element_blank(),
    strip.text = element_blank(),
    axis.title.x = element_blank()
  )
test_plots[[6]] <-
  test_plots[[6]] + labs(color = 'Sex') + theme(strip.background =  element_blank(),
                                                strip.text = element_blank(),
    axis.title.x = element_blank())
test_plots <- list(
  test_plots[[1]],
  test_plots[[6]],
  test_plots[[5]],
  test_plots[[2]],
  test_plots[[3]],
  test_plots[[4]]
)

exam_plots <- stats_table_plot_dat %>% 
   map(
    ~ ggplot(filter(.x, variable %!in% c('a1c', 'CHOL_all', 'flu')),
             aes(
      x = year,
      y = p,
      group = strata,
      color = strata
    )) +
      #geom_point() +
      geom_line(aes(x = year, y = model, group = strata, color = strata)) +
      geom_line(data = filter(overall_data, variable %!in% c('a1c', 'CHOL_all', 'flu')),
                  aes(x = year, y = model),
                  color = 'black',
                  linetype = 'dashed') +
      facet_wrap( ~ variable_text, ncol = 3) +
      theme_classic() +
      scale_y_continuous(
        limits = c(  0, 100),
        breaks = seq(0, 100, 25),
        labels = seq(0, 100, 25)
      ) +
      scale_x_continuous(
        limits = c(2008, 2020),
        breaks = seq(2008, 2020, 2),
        labels = seq(2008, 2020, 2)
      ) +
      scale_color_brewer(type = 'qual', palette = 'Set2') +
      theme(axis.text = element_text(color = 'black'),
            axis.text.x = element_text(angle = 45, hjust = 0.9, vjust = 0.95),
            legend.title = element_text(size = 8),
            legend.text = element_text(size = 7)) +
      labs(x = 'Year', y = 'Percentage')
  )

exam_plots[[1]] <- exam_plots[[1]] + labs(color = 'Age (grouped)') +
  theme(axis.title.x = element_blank())
exam_plots[[2]] <-
  exam_plots[[2]] + labs(color = 'Highest degree \nobtained') + theme(
    strip.background =  element_blank(),
    strip.text = element_blank(),
    axis.title.x = element_blank()
  )
exam_plots[[3]] <-
  exam_plots[[3]] + labs(color = "Insurance") + theme(
    strip.background =  element_blank(),
    strip.text = element_blank(),
    axis.title.x = element_blank()
  )
exam_plots[[4]] <-
  exam_plots[[4]] + labs(color = 'Poverty income ratio') + theme(
    strip.background =  element_blank(),
    strip.text = element_blank()
  )
exam_plots[[5]] <-
  exam_plots[[5]] + labs(color = 'Race & Ethnicity') + theme(
    strip.background =  element_blank(),
    strip.text = element_blank(),
    axis.title.x = element_blank()
  )
exam_plots[[6]] <-
  exam_plots[[6]] + labs(color = 'Sex') + theme(strip.background =  element_blank(),
                                                strip.text = element_blank(),
    axis.title.x = element_blank())
exam_plots <- list(
  exam_plots[[1]],
  exam_plots[[6]],
  exam_plots[[5]],
  exam_plots[[2]],
  exam_plots[[3]],
  exam_plots[[4]]
)

# all_exam_plots <- exam_plots %>% 
#   wrap_plots(nrow = 6) &
#   theme(plot.margin = margin(0.5, 0.5, 0.5, 0.5, 'pt'))
# all_test_plots <- test_plots %>% 
#   wrap_plots(nrow = 6) &
#   theme(plot.margin = margin(0.5, 0.5, 0.5, 0.5, 'pt'))
```

Youngest age group was consistently lower than all others for cholesterol testing and receiving flu vaccine, and eye exam and foot exam

Hispanic - lowest, although close for cholesterol and A1C (A1C until \~ 2018). And eye exam, but again real close

\< HS consistently lower on all exams and A1c tests Uninsured lower on everything

Flu - lots of increases. Hung et al found no change over 2007 - 2018 using NHIS data

Cholesterol - a few decreases

A1C - a few increases

Dentists - kind of all over the place; Aldosari 2022 found no difference in utilization between people with diabetes and without (so not a trend analysis). Luo et al 2018 found declines for people with diabetes from 2004 - 2014

Foot exams - mostly all flat, except older. PEraj 2019 found that Hispanic americans and asian-americans were significantly less likely to have a foot exam

Eye exams - mostly flat or decreasing, except uninsured. Chen 2020 found there were increases after Medicaid expansion but only for the first two years. Eppley 2019 found no change in adherance from 2005 to 2016 using NHANES

Mostly flat, some decreases.